### 2024-02-22 manuscript.qmd
working on regression-based artifact correction

---
title: "Replication of Gantman et al. 2024"
author: "219 2024"
format: html
---

```{r}
#| label: Load libraries
#| include: false

library(readr)
library(dplyr)
library(jsonlite)
library(ggplot2)
library(osfr)
```

```{r}
#| label: Download the data from the OSF
osf_retrieve_node("ngtha") %>%
  osf_ls_files() %>%
  osf_download(path="data/final", conflicts="skip")
```

```{r}
#| label: Load the behavioral data
#| include: false

behavioral_data <- read_csv('data/final/behavioral.csv')
```

```{r}
#| label: Filter out the subjects who got fashion words twice due to programming error
#| include: false

exclude_for_fashion_twice_subjects <- behavioral_data %>%
  filter(phase == "fashion", task == "response") %>%
  group_by(subject_id) %>%
  count() %>%
  filter(n == 600) %>%
  pull(subject_id)
```

```{r}
#| label: Flip responses for subject 01
#| include: false

behavioral_data <- behavioral_data %>%
  mutate(correct = ifelse(subject_id == "01", !correct, correct))
```

```{r}
#| label: Get subject non-response rate
#| include: false

subject_non_response_rate <- behavioral_data %>%
  filter(phase == "moral" | phase == "fashion") %>%
  filter(task == "response") %>%
  group_by(subject_id) %>%
  summarise(non_response_rate = mean(is.na(rt)))

exclude_for_non_response_subjects <- subject_non_response_rate %>%
  filter(non_response_rate > 0.5) %>%
  pull(subject_id)
```

```{r}
#| label: Compute subject non-word response rate
#| include: false

subject_non_word_response_rate <- behavioral_data %>%
  filter(phase == "moral" | phase == "fashion") %>%
  filter(task == "response") %>%
  group_by(subject_id) %>%
  mutate(non_word_response = ifelse(
    (is_word == FALSE & correct == TRUE) |
    (is_word == TRUE & correct == FALSE), 1, 0)) %>%
  summarize(non_word_response_rate = mean(non_word_response))

exclude_for_non_word_response_subjects <- subject_non_word_response_rate %>%
  filter(non_word_response_rate > 0.9) %>%
  pull(subject_id)
```

```{r}
#| label: Create filtered data set by applying exclusions
#| include: false

excluded_subjects <- c(exclude_for_fashion_twice_subjects, exclude_for_non_response_subjects, exclude_for_non_word_response_subjects)

behavioral_data_filtered <- behavioral_data %>%
  filter(!subject_id %in% excluded_subjects)
```

```{r}
#| label: Select relevant test data
#| include: false

behavioral_data_test_set <- behavioral_data_filtered %>%
  filter(phase == "moral" | phase == "fashion") %>%
  filter(task == "response") %>%
  select(subject_id, word_type, is_word, correct, rt) %>%
  mutate(word_category = ifelse(word_type %in% c("fashion", "moral"), "category", "non-category"),
         category = ifelse(word_type %in% c("fashion", "non-fashion"), "fashion", "moral"))
```

```{r}
#| label: Check accuracy
#| include: false

subject_level_accuracy <- behavioral_data_test_set %>%
  group_by(subject_id, word_category, category, is_word) %>%
  summarise(accuracy = mean(correct))
```

```{r}
#| label: fig-beh-pop-out

behavioral_pop_out_data <- subject_level_accuracy %>%
  filter(is_word == TRUE)

ggplot(behavioral_pop_out_data, aes(x=word_category, y=accuracy, group=subject_id))+
  geom_point()+
  geom_line()+
  #labs(x="Word Type", y="Accuracy")+
  facet_grid(.~category)+
  theme_bw()
```

```{r}
#| label: model the GEE
#| include: false

library(gee)

m <- gee(correct ~ word_category,
    id = subject_id,
    subset = is_word == TRUE & category == "moral",
    data = behavioral_data_test_set,
    family = binomial,
    corstr = "exchangeable")

summary(m)
```

```{r}
#| label: Load eeg data
#| include: false

eeg_data <- read_csv("data/final/eeg.csv", col_types = "cdcllcdlc") %>%
  filter(!subject %in% excluded_subjects)
```

```{r}
#| label: Flip subject 01s responses
#| include: false

eeg_data <- eeg_data %>%
  mutate(correct = ifelse(subject == "01", !correct, correct))
```

```{r}
#| label: Subject ERPs
#| include: false

subject_eeg <- eeg_data %>%
  filter(good_segment == TRUE) %>%
  filter(electrode %in% c("Cz", "Pz")) %>%
  filter(correct == TRUE) %>%
  group_by(subject, t, electrode, word_type, is_word) %>%
  summarize(v = mean(v), n=n()) %>%
  mutate(word_category = ifelse(word_type %in% c("fashion", "moral"), "category", "non-category"),
         category = ifelse(word_type %in% c("fashion", "non-fashion"), "fashion", "moral"))
```

```{r}
grand_average <- subject_eeg %>%
  group_by(t, electrode, word_category, category, is_word) %>%
  summarize(M = weighted.mean(v, n), SE = sd(v)/sqrt(n()))
```

```{r}
#| label: fig-grand-averages
#| fig-width: 10
#| fig-height: 8
ggplot(grand_average, aes(x=t, y=M, color=is_word, linetype=word_category)) +
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  geom_ribbon(aes(ymin=M-SE, ymax=M+SE, fill=is_word), alpha=0.2, color=NA)+
  geom_line()+
  scale_x_continuous(breaks=c(0,500, 1000))+
  facet_grid(electrode~category)+
  labs(x="Time since word onset (ms)", y="Voltage (uV)", color="Is word?", fill="Is word?", linetype="Word Category")+
  theme_minimal()+
  theme(panel.grid = element_blank())
```

```{r}
count_good_trials <- eeg_data %>%
  dplyr::filter(good_segment == TRUE) %>%
  dplyr::filter(electrode %in% c("Cz", "Pz")) %>%
  group_by(subject, electrode, word_type, is_word, event_id) %>%
  slice(1) %>%
  group_by(subject, electrode, word_type, is_word) %>%
  summarize(n = n())

count_good_trials
```