### 2024-03-14 manuscript.qmd
Update preprocessing to remove edge effects and filter segments with a bad channel.
Update manuscript.

---
title: "Replication of Gantman et al. 2024"
author: "219 2024"
format: html
---

```{r}
#| label: Load libraries
#| include: false

library(readr)
library(dplyr)
library(jsonlite)
library(ggplot2)
library(osfr)
library(gee)
library(tidyr)
```

# Introduction

This is a replication of the study by Gantman et al. (2014). Replications are important for *reasons*.

# Methods

```{r}
#| label: Download the data from the OSF
#| include: false
osf_retrieve_node("ngtha") %>%
  osf_ls_files() %>%
  osf_download(path="data/final", conflicts="skip")
```

```{r}
#| label: Load the behavioral data
#| include: false

behavioral_data <- read_csv('data/final/behavioral.csv')
```

## Participants

```{r}
#| label: Filter out the subjects who got fashion words twice due to programming error
#| include: false

exclude_for_fashion_twice_subjects <- behavioral_data %>%
  filter(phase == "fashion", task == "response") %>%
  group_by(subject_id) %>%
  count() %>%
  filter(n == 600) %>%
  pull(subject_id)
```

```{r}
#| label: Flip responses for subject 01, 36 moral words
#| include: false

behavioral_data <- behavioral_data %>%
  mutate(correct = ifelse(subject_id == "01", !correct, correct)) %>%
  mutate(correct = ifelse(subject_id == "36" & phase == "moral", !correct, correct))
```

```{r}
#| label: Get subject non-response rate
#| include: false

subject_non_response_rate <- behavioral_data %>%
  filter(phase == "moral" | phase == "fashion") %>%
  filter(task == "response") %>%
  group_by(subject_id) %>%
  summarise(non_response_rate = mean(is.na(rt)))

exclude_for_non_response_subjects <- subject_non_response_rate %>%
  filter(non_response_rate > 0.5) %>%
  pull(subject_id)
```

```{r}
#| label: Compute subject non-word response rate
#| include: false

subject_non_word_response_rate <- behavioral_data %>%
  filter(phase == "moral" | phase == "fashion") %>%
  filter(task == "response") %>%
  group_by(subject_id) %>%
  mutate(non_word_response = ifelse(
    (is_word == FALSE & correct == TRUE) |
    (is_word == TRUE & correct == FALSE), 1, 0)) %>%
  summarize(non_word_response_rate = mean(non_word_response))

exclude_for_non_word_response_subjects <- subject_non_word_response_rate %>%
  filter(non_word_response_rate > 0.9) %>%
  pull(subject_id)
```

```{r}
#| label: Create filtered data set by applying exclusions
#| include: false

excluded_subjects <- c(exclude_for_fashion_twice_subjects, exclude_for_non_response_subjects, exclude_for_non_word_response_subjects)

behavioral_data_filtered <- behavioral_data %>%
  filter(!subject_id %in% excluded_subjects)
```

```{r}
#| label: Count the number of subjects and subjects remaining
#| include: false

n_subjects <- behavioral_data %>% pull(subject_id) %>% unique() %>% length()

n_subjects_remaining <- behavioral_data_filtered %>% pull(subject_id) %>% unique() %>% length()
```

`r n_subjects` subjects were recruited for the study. After applying the exclusion criteria, `r n_subjects_remaining` subjects remain. `r length(exclude_for_fashion_twice_subjects)` subjects were excluded for receiving fashion words twice due to a technical error, `r length(exclude_for_non_response_subjects)` subjects were excluded for having a non-response rate greater than 50%, and `r length(exclude_for_non_word_response_subjects)` subjects were excluded for having a non-word response rate greater than 90%.

# Results

## Behavioral

```{r}
#| label: Select relevant test data
#| include: false

behavioral_data_test_set <- behavioral_data_filtered %>%
  filter(phase == "moral" | phase == "fashion") %>%
  filter(task == "response") %>%
  select(subject_id, word_type, is_word, correct, rt) %>%
  mutate(word_category = ifelse(word_type %in% c("fashion", "moral"), "category", "non-category"),
         category = ifelse(word_type %in% c("fashion", "non-fashion"), "fashion", "moral"))
```

```{r}
#| label: Check accuracy
#| include: false

subject_level_accuracy <- behavioral_data_test_set %>%
  group_by(subject_id, word_category, category, is_word) %>%
  summarise(accuracy = mean(correct), se=sd(correct)/sqrt(n()))
```

```{r}
#| label: fig-beh-pop-out
#| fig-cap: Accuracy for word trials in the fashion and moral conditions showed that performance was much higher than expected. We later discovered this was due to a difference in procedure between our study and the original study.

behavioral_pop_out_data <- subject_level_accuracy %>%
  filter(is_word == TRUE) %>%
  pivot_wider(names_from = word_category, values_from = c(accuracy, se))

ggplot(behavioral_pop_out_data, aes(x=`accuracy_non-category`, y=`accuracy_category`))+
  geom_point()+
  geom_linerange(aes(ymin=`accuracy_category`-`se_category`, ymax=`accuracy_category`+`se_category`), size=0.25)+
  geom_linerange(aes(xmin=`accuracy_non-category`-`se_non-category`, xmax=`accuracy_non-category`+`se_non-category`), size=0.25, alpha=0.3)+
  labs(x="Accuracy for non-category words", y="Accuracy for category words")+
  facet_grid(.~category, labeller = as_labeller(c("fashion"="Fashion Words", "moral"="Moral Words")))+ 
  geom_abline(intercept = 0, slope = 1, linetype="dashed")+
  coord_cartesian(xlim=c(0.5,1), ylim=c(0.5,1))+
  theme_bw(base_size=12)+
  theme(aspect.ratio=1, panel.grid=element_blank())
```

```{r}
#| label: Models for pop out effects
#| include: false

behavioral_model_moral_pop_out <- gee(correct ~ word_category,
    id = subject_id,
    subset = is_word == TRUE & category == "moral",
    data = behavioral_data_test_set,
    family = binomial,
    corstr = "exchangeable") %>% summary()

coefficient_moral_pop_out <- behavioral_model_moral_pop_out$coefficients[2, 'Estimate']
coefficient_moral_pop_out_se <- behavioral_model_moral_pop_out$coefficients[2, 'Robust S.E.']
z_moral_pop_out <- behavioral_model_moral_pop_out$coefficients[2, 'Robust z']
p_moral_pop_out <- z_moral_pop_out %>% pnorm(lower.tail = FALSE)

fashion_model_moral_pop_out <- gee(correct ~ word_category,
    id = subject_id,
    subset = is_word == TRUE & category == "fashion",
    data = behavioral_data_test_set,
    family = binomial,
    corstr = "exchangeable") %>% summary()

coefficient_fashion_pop_out <- fashion_model_moral_pop_out$coefficients[2, 'Estimate']
coefficient_fashion_pop_out_se <- fashion_model_moral_pop_out$coefficients[2, 'Robust S.E.']
z_fashion_pop_out <- fashion_model_moral_pop_out$coefficients[2, 'Robust z']
p_fashion_pop_out <- z_fashion_pop_out %>% pnorm(lower.tail = FALSE)
```

Accuracy for moral words was not significantly different between category and non-category words, $b = `r coefficient_moral_pop_out`$, $SE = `r coefficient_moral_pop_out_se`$, $z = `r z_moral_pop_out`$, $p = `r p_moral_pop_out`$. Accuracy for fashion words was significantly different between category and non-category words, $b = `r coefficient_fashion_pop_out`$, $SE = `r coefficient_fashion_pop_out_se`$, $z = `r z_fashion_pop_out`$, $p = `r p_fashion_pop_out`$.

```{r}
#| label: Compute response time differences for words
#| include: false

subject_level_rt <- behavioral_data_test_set %>%
  filter(correct == TRUE) %>%
  group_by(subject_id, word_category, category, is_word) %>%
  summarize(rt = mean(rt, na.rm = TRUE))
```

```{r}
#| include: false

behavioral_pop_out_data <- subject_level_rt %>%
  filter(is_word == TRUE)

ggplot(behavioral_pop_out_data, aes(x=word_category, y=rt, group=subject_id))+
  geom_point()+
  geom_line()+
  #labs(x="Word Type", y="Accuracy")+
  facet_grid(.~category)+
  theme_bw()
```

```{r}
#| include: false

m <- gee(rt ~ word_category,
    id = subject_id,
    subset = is_word == TRUE & correct == TRUE & category == "moral",
    data = behavioral_data_test_set,
    family = gaussian,
    corstr = "exchangeable")

summary(m)
```

## EEG

```{r}
#| label: Load eeg data
#| include: false

eeg_data <- read_csv("data/final/eeg.csv", col_types = "cdcllcdc") %>%
  filter(!subject %in% excluded_subjects) %>%
  filter(subject != "34") # really few usable segments
```

```{r}
#| label: Flip subject 01s responses
#| include: false

eeg_data <- eeg_data %>%
  mutate(correct = ifelse(subject == "01", !correct, correct)) %>%
  mutate(correct = ifelse(subject == "36" & word_type == "moral", !correct, correct))
```

```{r}
#| label: Subject ERPs
#| include: false

subject_eeg <- eeg_data %>%
  filter(electrode %in% c("Cz", "Pz")) %>%
  filter(correct == TRUE) %>%
  group_by(subject, t, electrode, word_type, is_word) %>%
  summarize(v = mean(v), n=n()) %>%
  mutate(word_category = ifelse(word_type %in% c("fashion", "moral"), "category", "non-category"),
         category = ifelse(word_type %in% c("fashion", "non-fashion"), "fashion", "moral"))
```

```{r}
grand_average <- subject_eeg %>%
  group_by(t, electrode, word_category, category, is_word) %>%
  summarize(M = mean(v), SE = sd(v)/sqrt(n()))
```

```{r}
#| label: fig-grand-averages
#| fig-width: 10
#| fig-height: 8
ggplot(grand_average, aes(x=t, y=M, color=is_word, linetype=word_category)) +
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  geom_ribbon(aes(ymin=M-SE, ymax=M+SE, fill=is_word), alpha=0.2, color=NA)+
  geom_line()+
  scale_x_continuous(breaks=c(0,500, 1000))+
  facet_grid(electrode~category)+
  labs(x="Time since word onset (ms)", y="Voltage (uV)", color="Is word?", fill="Is word?", linetype="Word Category")+
  theme_minimal()+
  theme(panel.grid = element_blank())
```

```{r}
#| label: Count the number of good segments per subject
#| include: false
count_good_trials <- eeg_data %>%
  filter(t==0, electrode=="Cz") %>%
  group_by(subject) %>%
  summarize(n=n()) %>%
  mutate(percent = (n/600)*100)

mean_good_trials <- mean(count_good_trials$n)
sd_good_trials <- sd(count_good_trials$n)

```

```{r}
#| label: Define time windows
#| include: false
p2.window.start <- 200
p2.window.end <- 250
n2.window.start <- 250
n2.window.end <- 350
p3.window.start <- 350
p3.window.end <- 600
lpp.window.start <- 600
lpp.window.end <- 800

# vectorized function for indicating time window
get_time_window <- function(t){
  ifelse(t >= p2.window.start & t < p2.window.end, "p2",
  ifelse(t >= n2.window.start & t < n2.window.end, "n2",
  ifelse(t >= p3.window.start & t < p3.window.end, "p3",
  ifelse(t >= lpp.window.start & t < lpp.window.end, "lpp",
  "other"))))
}
```

```{r}
```

```{r}
#| label: Calculate mean voltage in all time windows for each trial
#| include: false

time_window_data <- eeg_data %>%
  filter(electrode %in% c("Cz", "Pz")) %>%
  mutate(erp_component = get_time_window(t)) %>%
  filter(erp_component != "other") %>%
  group_by(subject, electrode, word_type, is_word, erp_component) %>%
  summarize(v = mean(v)) %>%
  mutate(word_category = ifelse(word_type %in% c("fashion", "moral"), "category", "non-category"),
         category = ifelse(word_type %in% c("fashion", "non-fashion"), "fashion", "moral"))
```

```{r}
#| label: Run GEE models for all time windows, fashion data pop-out
#| include: false

p2_model <- gee(v ~ word_category,
    id = subject,
    subset = erp_component == "p2" & category == "fashion" & is_word == TRUE & electrode == "Cz",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

p2_model$coefficients

n2_model <- gee(v ~ word_category,
    id = subject,
    subset = erp_component == "n2" & category == "fashion"& is_word == TRUE & electrode == "Cz",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

n2_model$coefficients

p3_model <- gee(v ~ word_category,
    id = subject,
    subset = erp_component == "p3" & category == "fashion"& is_word == TRUE & electrode == "Cz",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

p3_model$coefficients

lpp_model <- gee(v ~ word_category,
    id = subject,
    subset = erp_component == "lpp" & category == "fashion"& is_word == TRUE & electrode == "Cz",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

lpp_model$coefficients
```
```{r}
#| label: Run GEE models for all time windows, moral data pop-out
#| include: false

p2_model <- gee(v ~ word_category,
    id = subject,
    subset = erp_component == "p2" & category == "moral" & is_word == TRUE & electrode == "Cz",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

p2_model$coefficients

n2_model <- gee(v ~ word_category,
    id = subject,
    subset = erp_component == "n2" & category == "moral" & is_word == TRUE & electrode == "Cz",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

n2_model$coefficients

p3_model <- gee(v ~ word_category,
    id = subject,
    subset = erp_component == "p3" & category == "moral"& is_word == TRUE & electrode == "Cz",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

p3_model$coefficients

lpp_model <- gee(v ~ word_category,
    id = subject,
    subset = erp_component == "lpp" & category == "moral"& is_word == TRUE & electrode == "Cz",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

lpp_model$coefficients
```

```{r}
#| label: Run GEE models for all time windows, moral data word non-word
#| include: false

p2_model <- gee(v ~ is_word,
    id = subject,
    subset = erp_component == "p2" & category == "moral",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

p2_model$coefficients

n2_model <- gee(v ~ is_word,
    id = subject,
    subset = erp_component == "n2" & category == "moral",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

n2_model$coefficients

p3_model <- gee(v ~ is_word,
    id = subject,
    subset = erp_component == "p3" & category == "moral",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

p3_model$coefficients

lpp_model <- gee(v ~ is_word,
    id = subject,
    subset = erp_component == "lpp" & category == "moral",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

lpp_model$coefficients

```

```{r}
#| label: Run GEE models for all time windows, fashion data word non-word
#| include: false

p2_model <- gee(v ~ is_word,
    id = subject,
    subset = erp_component == "p2" & category == "fashion",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

p2_model$coefficients

n2_model <- gee(v ~ is_word,
    id = subject,
    subset = erp_component == "n2" & category == "fashion",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

n2_model$coefficients

p3_model <- gee(v ~ is_word,
    id = subject,
    subset = erp_component == "p3" & category == "fashion",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

p3_model$coefficients

lpp_model <- gee(v ~ is_word,
    id = subject,
    subset = erp_component == "lpp" & category == "fashion",
    data = time_window_data,
    family = gaussian,
    corstr = "exchangeable") %>% summary()

lpp_model$coefficients

```